Manuscript

### 2019-01-24 manuscript.Rmd

We derived our clinical associations from a new project conducted by the Variant Interpretation for Cancer Consortium (VICC; https://cancervariants.org/), a Driver Project for the Global Alliance for Genomics and Health (GA4GH) [@theglobalallianceforgenomicsandhealth2016]. The VICC has curated annotations of known cancer variants at varying levels of evidence from multiple resources [@wagner2018]: the Precision Medicine Knowledgebase [@huang2016], MolecularMatch (https://www.molecularmatch.com/index.html), OncoKB [@chakravarty2017], Jackson Labs Clinical Knowledgebase [@patterson2016], Clinical Interpretations of Variance in Cancers (CIViC) [@griffith2017], and the Cancer Genome Interpreter (CGI) [@tamborero2018]. These harmonized variants are hosted on an ElasticSearch (Kibana v6.0) platform called Genotype to Phenotype (G2P, https://search.cancervariants.org/#*) that allows users to query and filter aggregated evidence from the various databases listed above as well as the GA4GH beacon service (http://beacon-network.org/#/), which yields access to genetic mutation data from over 200 datasets as of 2016 [@theglobalallianceforgenomicsandhealth2016]. For drug screen datasets, we focus on three well-known, large-scale drug projects: the CCLE and GDSC, which were mentioned above, and the Center for the Science of Therapeutics at the Broad Institute’s Cancer Therapeutics Response Portal (CTRP) dataset [@basu2013; @rees2016; @seashore-ludlow2015]. We also analyzed the results from a large-scale CRISPR/Cas9 screen conducted by the Broad Institute’s Cancer Dependency Map project (DepMap) using the Avana knockout library [@doench2016a; @meyers2017], which will address whether gene essentiality screens via CRISPR yield results comparable to functional drug screens.

@@

dg2intext_converter <- function(drug_gene) {
  # Convert drug-gene strings to have italic gene name
  d_g <- strsplit(drug_gene, split = "_")
  dg <- paste0(d_g[[1]][1], "-*", d_g[[1]][2], "*")
  return(dg)
}
```

@@

```{r drug-data, echo=FALSE, include=FALSE}
ccle_data <- readRDS("./data_munging/rds/ccle_data_18Q4_g2pgenes.rds")
ccle_data_ptmuts <- filter(ccle_data, Reference_Allele_Length == 1 | Reference_Allele_Length == 0 | is.na(Reference_Allele_Length))
ccle_data_ptmuts$Keep_PtMut_Tests <- ifelse(ccle_data_ptmuts$Reference_Allele_Length == 1 | is.na(ccle_data_ptmuts$Reference_Allele_Length), TRUE, ifelse(ccle_data_ptmuts$Tumor_Seq_Allele1_Length == 1, TRUE, FALSE))
ccle_data_ptmuts <- filter(ccle_data_ptmuts, Keep_PtMut_Tests == TRUE)
ccle_data_ptmuts$Dataset <- "CCLE"
ccle_data_ptmuts$Hugo_Symbol <- ordered(ccle_data_ptmuts$Hugo_Symbol, levels = g2p_genes)

ctrp_data <- readRDS("./data_munging/rds/ctrp_data_18Q4_g2pgenes.rds")
ctrp_data_ptmuts <- filter(ctrp_data, Reference_Allele_Length == 1 | Reference_Allele_Length == 0 | is.na(Reference_Allele_Length))
ctrp_data_ptmuts$Keep_PtMut_Tests <- ifelse(ctrp_data_ptmuts$Reference_Allele_Length == 1 | is.na(ctrp_data_ptmuts$Reference_Allele_Length), TRUE, ifelse(ctrp_data_ptmuts$Tumor_Seq_Allele1_Length == 1, TRUE, FALSE))
ctrp_data_ptmuts <- filter(ctrp_data_ptmuts, Keep_PtMut_Tests == TRUE)
ctrp_data_ptmuts$Dataset <- "CTRP"
ctrp_data_ptmuts$Hugo_Symbol <- ordered(ctrp_data_ptmuts$Hugo_Symbol, levels = g2p_genes)

gdsc_data <- readRDS("./data_munging/rds/gdsc_data_18Q4_g2pgenes.rds")
gdsc_data_ptmuts <- filter(gdsc_data, Reference_Allele_Length == 1 | Reference_Allele_Length == 0 | is.na(Reference_Allele_Length))
gdsc_data_ptmuts$Keep_PtMut_Tests <- ifelse(gdsc_data_ptmuts$Reference_Allele_Length == 1 | is.na(gdsc_data_ptmuts$Reference_Allele_Length), TRUE, ifelse(gdsc_data_ptmuts$Tumor_Seq_Allele1_Length == 1, TRUE, FALSE))
gdsc_data_ptmuts <- filter(gdsc_data_ptmuts, Keep_PtMut_Tests == TRUE)
gdsc_data_ptmuts$Dataset <- "GDSC"
gdsc_data_ptmuts$Hugo_Symbol <- ordered(gdsc_data_ptmuts$Hugo_Symbol, levels = g2p_genes)

@@

After filtering, `r nrow(g2p_filt)` level-A G2P interpretations were carried forward in subsequent analyses, constituting `r length(g2p_drugs)` unique drugs and `r length(g2p_genes)` unique genes in `r length(g2p_druggenes)` distinct level-A combinations (*Figure 1A*). The number of level-A interpretations per drug ranged from `r range(g2p_drug_tally$n)[1]` to `r range(g2p_drug_tally$n)[2]` ($M =$ `r printnum(mean(g2p_drug_tally$n))`, $SD =$ `r printnum(sd(g2p_drug_tally$n))`), `r range(g2p_gene_tally$n)[1]` to `r range(g2p_gene_tally$n)[2]` ($M =$ `r printnum(mean(g2p_gene_tally$n))`, $SD =$ `r printnum(sd(g2p_gene_tally$n))`) per gene, and for each level-A drug-gene combination, the range was `r range(g2p_druggene_tally$n)[1]` to `r range(g2p_druggene_tally$n)[2]` ($M =$ `r printnum(mean(g2p_druggene_tally$n))`, $SD =$ `r printnum(sd(g2p_druggene_tally$n))`). These unique drugs, genes, and combinations are hereafter referred to as G2P drugs, G2P genes, and G2P associations. Filtering our screening datasets by these clinical criteria yields panels of potentially clinically relevant omic features. Of the `r length(g2p_drugs)` unique G2P drugs, `r nrow(drug_summ[drug_summ$Dataset_Count >= 1,])` of them were screened in at least one of the drug screen datasets, `r nrow(drug_summ[drug_summ$Dataset_Count >= 2,])` drugs were screened in at least two of the three datasets, and only `r nrow(drug_summ[drug_summ$Dataset_Count >= 3,])` were screened in all three datasets. The remaining `r nrow(drug_summ[drug_summ$Dataset_Count == 0,])` un-screened G2P drugs suggests an incompleteness of these drug screens, though these projects frequently release new screening data that may include these drugs in future releases.

Using cell-line-specific point mutation annotations from the CCLE, we compared z-score-transformed area under the dose-response curve (AUC) distributions between mutant and wildtype cell lines for all G2P associations tested in the CCLE, CTRP, and GDSC drug screens. Of the `r length(g2p_druggenes)` G2P associations, data was available for `r length(ccle_signif_g2p$Drug_Gene)` associations in CCLE, `r length(ctrp_signif_g2p$Drug_Gene)` in CTRP, and `r length(gdsc_signif_g2p$Drug_Gene)` in GDSC. Controlling for a false discovery rate (FDR) < 5%, CCLE yielded no significant correlations (Wilcoxon rank-sum tests, `r ccle_signif_g2p_exp`), though the most significant G2P association was `r dg2intext_converter(ccle_signif_g2p$Drug_Gene[1])` (`r e2exp_converter(ccle_signif_g2p$p.adj[1], p_type = "padj")`). In CTRP, AUC from `r nrow(filter(ctrp_signif_g2p[!is.na(ctrp_signif_g2p$p.signif.adj),]))` drug-gene associations correlated with the gene's mutation status (`r ctrp_signif_g2p_exp`): `r dg2intext_converter(ctrp_signif_g2p$Drug_Gene[1])` (`r e2exp_converter(ctrp_signif_g2p$p.adj[1], p_type = "padj")`), `r dg2intext_converter(ctrp_signif_g2p$Drug_Gene[2])` (`r e2exp_converter(ctrp_signif_g2p$p.adj[2], p_type = "padj")`), and `r dg2intext_converter(ctrp_signif_g2p$Drug_Gene[3])` (`r e2exp_converter(ctrp_signif_g2p$p.adj[3], p_type = "padj")`). `r str_ucfirst(dg2intext_converter(gdsc_signif_g2p$Drug_Gene[1]))` (`r e2exp_converter(gdsc_signif_g2p$p.adj[1], p_type = "padj")`) and `r dg2intext_converter(gdsc_signif_g2p$Drug_Gene[2])` (`r e2exp_converter(gdsc_signif_g2p$p.adj[2], p_type = "padj")`) were significant in GDSC (`r gdsc_signif_g2p_exp`).

@@

To test for potentially novel drug-gene associations, we conducted similar Wilcoxon rank-sum tests for all pairwise combinations of G2P genes and drugs tested in the three drug screens (*Figures 1B and 2A*). At an unadjusted $\alpha = 0.05$, the tests revealed that in CCLE, AUC significantly correlated with mutation status for `r nrow(filter(ccle_signif_g2p_grid, p < 0.05))` out of `r nrow(ccle_signif_g2p_grid)` (`r printnum(nrow(filter(ccle_signif_g2p_grid, p < 0.05)) / nrow(ccle_signif_g2p_grid) * 100)`%) tested drug-gene combinations were significantly different, `r nrow(filter(ctrp_signif_g2p_grid, p < 0.05))` out of `r nrow(ctrp_signif_g2p_grid)` (`r printnum(nrow(filter(ctrp_signif_g2p_grid, p < 0.05)) / nrow(ctrp_signif_g2p_grid) * 100)`%) were significantly different for CTRP, and `r nrow(filter(gdsc_signif_g2p_grid, p < 0.05))` out of `r nrow(gdsc_signif_g2p_grid)` (`r printnum(nrow(filter(gdsc_signif_g2p_grid, p < 0.05)) / nrow(gdsc_signif_g2p_grid) * 100)`%) for GDSC.

Controlling for a FDR < 5%, `r dg2intext_converter(ccle_signif_g2p_grid$Drug_Gene[1])` (`r e2exp_converter(ccle_signif_g2p_grid$p.adj[1], p_type = "padj")`) was the only significant drug-gene combination tested in CCLE (`r ccle_signif_g2p_grid_exp`). CTRP (`r ctrp_signif_g2p_grid_exp`) yielded `r nrow(ctrp_signif_g2p_grid[!is.na(ctrp_signif_g2p_grid$p.signif.adj),])` significant combinations, one of which was selumetinib-*BRAF*. There were `r nrow(gdsc_signif_g2p_grid[!is.na(gdsc_signif_g2p_grid$p.signif.adj),])` significant G2P associations in the GDSC dataset (`r gdsc_signif_g2p_grid_exp`). Of all the significant correlations between mutation status and AUC across datasets, the only genes represented were *KRAS* and *BRAF* (Figure\ \@ref(fig:drug-venn)). Of the `r length(drug_grid_shared_dg)` significant drug-gene associations across datasets, `r length(intersect(drug_grid_shared_dg, g2p_druggenes))` were G2P associations. The remaining `r length(drug_grid_shared_dg) - length(intersect(drug_grid_shared_dg, g2p_druggenes))` associations were `r dg2intext_converter(setdiff(drug_grid_shared_dg, g2p_druggenes)[1])`, `r dg2intext_converter(setdiff(drug_grid_shared_dg, g2p_druggenes)[2])`, and `r dg2intext_converter(setdiff(drug_grid_shared_dg, g2p_druggenes)[3])`.

@@

(ref:drug-venn) Overlap of drug-gene associations drawn from all pairwise combinations of G2P drugs and G2P genes whose drug effects significantly correlated with (**A**) mutation status, (**B**) gene expression, and (**C**) copy number across CCLE, CTRP, and GDSC.

@@

Prior to correction for an FDR of 5%, no DGLs were identified in the CCLE screen when using mutation status as a correlate ($n =$ `r nrow(ccle_signif_g2p_lineage)`, `r length(unique(ccle_data_g2p_dgl$Drug_Gene_Lin)) - nrow(ccle_signif_g2p_lineage)` DGLs were not analyzed due to lack of mutation annotations in *UGT1A1* in "cancer" lineages), but CTRP ($n =$ `r nrow(ctrp_signif_g2p_lineage)`) and GDSC ($n =$ `r nrow(gdsc_signif_g2p_lineage)`) identified `r nrow(filter(ctrp_signif_g2p_lineage, !is.na(p.signif)))` and `r nrow(filter(gdsc_signif_g2p_lineage, !is.na(p.signif)))`, respectively, in skin cancer, lung cancer, and cancer. After Benjamini-Hochberg FDR correction, none of the datasets yielded significant DGL interactions.

> **(???) KS test results**

@@

(ref:ksplot) Comparison of Spearman correlation coefficients ($r_s$) derived from copy number (log$_2$ ratio), gene expression (RPKM), and AUC z-scores from G2P drug-gene associations. Coefficients were grouped by whether or not the associated lineage for each comparison was in a G2P drug-gene-lineage association or not. Distributions of these grouped coefficients were compared using Kolmogorov-Smirnov tests.

@@

```{r crispr-data, echo = FALSE, include = FALSE, warning = FALSE, message = FALSE}
crispr_data <- readRDS("./../htscreens_giant_files/crispr_data_18Q4.rds")
crispr_data_ptmuts <- filter(crispr_data, Reference_Allele_Length == 1 | Reference_Allele_Length == 0 | is.na(Reference_Allele_Length))
crispr_data_ptmuts$Keep_PtMut_Tests <- ifelse(crispr_data_ptmuts$Reference_Allele_Length == 1 | is.na(crispr_data_ptmuts$Reference_Allele_Length), TRUE, ifelse(crispr_data_ptmuts$Tumor_Seq_Allele1_Length == 1, TRUE, FALSE))
crispr_data_ptmuts <- filter(crispr_data_ptmuts, Keep_PtMut_Tests == TRUE)

# Filter CRISPR for G2P genes in G2P
crispr_data_g2p <- filter(crispr_data_ptmuts, Hugo_Symbol %in% g2p_genes)
crispr_data_g2p$Hugo_Symbol <- factor(crispr_data_g2p$Hugo_Symbol, levels = g2p_genes)
crispr_data_g2p$Mutation_Status_Nonsilent <- factor(crispr_data_g2p$Mutation_Status_Nonsilent, levels = c("Mutant", "Wildtype"))

@@

(ref:crispr-venn) Overlap of genes whose gene expression and/or copy number significantly correlated with CERES score.

@@

When we don’t account for lineage, Wilcoxon rank-sum tests (`r crispr_signif_g2p_gene_exp`, FDR < 5%) revealed that mutations in `r nrow(crispr_signif_g2p_gene_sig)` genes significantly correlated with CERES score: *`r as.character(crispr_signif_g2p_gene_sig$Hugo_Symbol)[1]`* (`r e2exp_converter(crispr_signif_g2p_gene_sig$p.adj[1], p_type = "padj")`) and *`r as.character(crispr_signif_g2p_gene_sig$Hugo_Symbol)[2]`* (`r e2exp_converter(crispr_signif_g2p_gene_sig$p.adj[2], p_type = "padj")`). Gene expression (RPKM) of `r nrow(crispr_spearman_res_GE_sig)` out of `r nrow(crispr_spearman_res_GE)` (`r printnum(nrow(crispr_spearman_res_GE_sig) / nrow(crispr_spearman_res_GE) * 100)`%) G2P genes significantly correlated with CERES score, as did copy number (log2-ratio) for `r nrow(crispr_spearman_res_CN_sig)` out of `r nrow(crispr_spearman_res_CN)` (`r printnum(nrow(crispr_spearman_res_CN_sig) / nrow(crispr_spearman_res_CN) * 100)`%) G2P genes (Figure\ \@ref(fig:crispr-venn)).

For the purpose of comparison to the drug screen data, if we treat CRISPR knockdown like a small-molecule drug, then these results are more comprehensive than the results from CCLE, CTRP, and GDSC (Table 2).

- Lineage?

@@

In a lineage-agnostic context of the CRISPR screen, ERBB2 gene essentiality score had a significant negative correlation with ERBB2 gene expression ($p < 10^{-12}$, $r = -0.31$) and ERBB2 copy number ($p < 10^{-15}$, $r = -0.34$) (*Figure ##2*). Similarly, MET gene essentiality also correlated with both MET gene expression ($p < 10^{-12}$, $r = -0.31$) and copy number ($p < 10^{-8}$, $r = -0.25$). Notably, BRCA1 also significantly correlated with both gene expression ($p < 0.001$, $r = 0.16$) and copy number ($p < 0.01$, $r = 0.14$). BRCA2 significantly correlated with all three of the genomic features tested: gene expression ($p < 0.05$, $r = 0.10$), copy number ($p < 10^{-7}$, $r = 0.23$), and mutation status ($p < 0.001$).

@@

**Should we do this?**

@@

For computing Spearman correlation coefficients, we used the `cor.test` base R package. If there were 3 or fewer data points, then the correlation was not done and a value of `NA` was supplied.

>>>>

To curate an evidence-based list of these clinical biomarkers, we utilized a new oncological data aggregation project conducted by the Variant Interpretation for Cancer Consortium (VICC; https://cancervariants.org/), a Driver Project for the Global Alliance for Genomics and Health (GA4GH) [@theglobalallianceforgenomicsandhealth2016]. The VICC has curated annotations of known cancer variants at varying levels of evidence from multiple resources [@wagner2018]: the Precision Medicine Knowledgebase [@huang2016], MolecularMatch (https://www.molecularmatch.com/index.html), OncoKB [@chakravarty2017], Jackson Labs Clinical Knowledgebase [@patterson2016], Clinical Interpretations of Variance in Cancers (CIViC) [@griffith2017], and the Cancer Genome Interpreter (CGI) [@tamborero2018]. These harmonized variants are hosted on an ElasticSearch (Kibana v6.0) platform called Genotype to Phenotype (G2P, https://search.cancervariants.org/#*) that allows users to query and filter aggregated evidence from the various databases listed above as well as the GA4GH beacon service (http://beacon-network.org/#/), which yields access to genetic mutation data from over 200 datasets as of 2016 [@theglobalallianceforgenomicsandhealth2016]. For drug screen datasets, we focus on three well-known, large-scale drug projects: the CCLE and GDSC, which were mentioned above, and the Center for the Science of Therapeutics at the Broad Institute’s Cancer Therapeutics Response Portal (CTRP) dataset [@basu2013; @rees2016; @seashore-ludlow2015]. We also analyzed the results from a large-scale CRISPR/Cas9 screen conducted by the Broad Institute’s Cancer Dependency Map project (DepMap) using the Avana knockout library [@doench2016a; @meyers2017], which will address whether gene essentiality screens via CRISPR yield results comparable to functional drug screens.

@@

drugstring2intext_converter <- function(string, type) {
  # Convert drug-gene/drug-gene-lineage strings to have italic gene name
  if(type == "dg") {
    d_g <- strsplit(string, split = "_")
    text <- paste0(d_g[[1]][1], "-*", d_g[[1]][2], "*")
  }
  if(type == "dgl") {
    d_g_l <- strsplit(string, split = "_")
    text <- paste0(d_g_l[[1]][1], "-*", d_g_l[[1]][2], "*-", d_g_l[[1]][3])
  }  
  return(text)
}
```

@@

```{r drug-data, echo=FALSE, include=FALSE}
ccle_data_ptmuts <- readRDS("./data_munging/rds/ccle_data_18Q4_g2pgenes_ptmuts.rds")
ctrp_data_ptmuts <- readRDS("./data_munging/rds/ctrp_data_18Q4_g2pgenes_ptmuts.rds")
gdsc_data_ptmuts <- readRDS("./data_munging/rds/gdsc_data_18Q4_g2pgenes_ptmuts.rds")

@@

After filtering, `r nrow(g2p_filt)` level-A G2P interpretations were carried forward in subsequent analyses, constituting `r length(g2p_drugs)` unique drugs and `r length(g2p_genes)` unique genes in `r length(g2p_druggenes)` distinct level-A combinations (*Figure 1A: Pink heatmap*). The number of level-A interpretations per drug ranged from `r range(g2p_drug_tally$n)[1]` to `r range(g2p_drug_tally$n)[2]` ($M =$ `r printnum(mean(g2p_drug_tally$n))`, $SD =$ `r printnum(sd(g2p_drug_tally$n))`), `r range(g2p_gene_tally$n)[1]` to `r range(g2p_gene_tally$n)[2]` ($M =$ `r printnum(mean(g2p_gene_tally$n))`, $SD =$ `r printnum(sd(g2p_gene_tally$n))`) per gene, and for each level-A drug-gene combination, the range was `r range(g2p_druggene_tally$n)[1]` to `r range(g2p_druggene_tally$n)[2]` ($M =$ `r printnum(mean(g2p_druggene_tally$n))`, $SD =$ `r printnum(sd(g2p_druggene_tally$n))`). These unique drugs, genes, and combinations are hereafter referred to as G2P drugs, G2P genes, and G2P associations. Filtering our screening datasets by these clinical criteria yields panels of potentially clinically relevant omic features. Of the `r length(g2p_drugs)` unique G2P drugs, `r nrow(drug_summ[drug_summ$Dataset_Count >= 1,])` of them were screened in at least one of the drug screen datasets, `r nrow(drug_summ[drug_summ$Dataset_Count >= 2,])` drugs were screened in at least two of the three datasets, and only `r nrow(drug_summ[drug_summ$Dataset_Count >= 3,])` were screened in all three datasets. The remaining `r nrow(drug_summ[drug_summ$Dataset_Count == 0,])` un-screened G2P drugs suggests an incompleteness of these drug screens, though these projects frequently release new screening data that may include these drugs in future releases.

Using cell-line-specific point mutation annotations from the CCLE, we compared z-score-transformed area under the dose-response curve (AUC) distributions between mutant and wildtype cell lines for all G2P associations tested in the CCLE, CTRP, and GDSC drug screens. Of the `r length(g2p_druggenes)` G2P associations, data was available for `r length(ccle_signif_g2p$Drug_Gene)` associations in CCLE, `r length(ctrp_signif_g2p$Drug_Gene)` in CTRP, and `r length(gdsc_signif_g2p$Drug_Gene)` in GDSC. Controlling for a false discovery rate (FDR) < 5%, CCLE yielded no significant correlations (Wilcoxon rank-sum tests, `r ccle_signif_g2p_exp`), though the most significant G2P association was `r drugstring2intext_converter(ccle_signif_g2p$Drug_Gene[1], "dg")` (`r e2exp_converter(ccle_signif_g2p$p.adj[1], p_type = "padj")`). In CTRP, AUC from `r nrow(filter(ctrp_signif_g2p[!is.na(ctrp_signif_g2p$p.signif.adj),]))` drug-gene associations correlated with the gene's mutation status (`r ctrp_signif_g2p_exp`): `r drugstring2intext_converter(ctrp_signif_g2p$Drug_Gene[1], "dg")` (`r e2exp_converter(ctrp_signif_g2p$p.adj[1], p_type = "padj")`), `r drugstring2intext_converter(ctrp_signif_g2p$Drug_Gene[2], "dg")` (`r e2exp_converter(ctrp_signif_g2p$p.adj[2], p_type = "padj")`), and `r drugstring2intext_converter(ctrp_signif_g2p$Drug_Gene[3], "dg")` (`r e2exp_converter(ctrp_signif_g2p$p.adj[3], p_type = "padj")`). `r str_ucfirst(drugstring2intext_converter(gdsc_signif_g2p$Drug_Gene[1], "dg"))` (`r e2exp_converter(gdsc_signif_g2p$p.adj[1], p_type = "padj")`) was the only significant association in GDSC (`r gdsc_signif_g2p_exp`).

@@

To test for potentially novel drug-gene associations, we conducted similar Wilcoxon rank-sum tests for all pairwise combinations of G2P genes and drugs tested in the three drug screens (*Figures 1B: Green heatmaps*). At an unadjusted $\alpha = 0.05$, the tests revealed that in CCLE, AUC significantly correlated with mutation status for `r nrow(filter(ccle_signif_g2p_grid, p < 0.05))` out of `r nrow(ccle_signif_g2p_grid)` (`r printnum(nrow(filter(ccle_signif_g2p_grid, p < 0.05)) / nrow(ccle_signif_g2p_grid) * 100)`%) tested drug-gene combinations were significantly different, `r nrow(filter(ctrp_signif_g2p_grid, p < 0.05))` out of `r nrow(ctrp_signif_g2p_grid)` (`r printnum(nrow(filter(ctrp_signif_g2p_grid, p < 0.05)) / nrow(ctrp_signif_g2p_grid) * 100)`%) were significantly different for CTRP, and `r nrow(filter(gdsc_signif_g2p_grid, p < 0.05))` out of `r nrow(gdsc_signif_g2p_grid)` (`r printnum(nrow(filter(gdsc_signif_g2p_grid, p < 0.05)) / nrow(gdsc_signif_g2p_grid) * 100)`%) for GDSC.

Controlling for a FDR < 5%, `r drugstring2intext_converter(ccle_signif_g2p_grid$Drug_Gene[1], "dg")` (`r e2exp_converter(ccle_signif_g2p_grid$p.adj[1], p_type = "padj")`) was the only significant drug-gene combination tested in CCLE (`r ccle_signif_g2p_grid_exp`). CTRP (`r ctrp_signif_g2p_grid_exp`) yielded `r nrow(ctrp_signif_g2p_grid[!is.na(ctrp_signif_g2p_grid$p.signif.adj),])` significant combinations, one of which was selumetinib-*BRAF*. There were `r nrow(gdsc_signif_g2p_grid[!is.na(gdsc_signif_g2p_grid$p.signif.adj),])` significant G2P associations in the GDSC dataset (`r gdsc_signif_g2p_grid_exp`). Of all the significant correlations between mutation status and AUC across datasets, the only genes represented were *KRAS* and *BRAF* (Figure\ \@ref(fig:drug-venn)). Of the `r length(drug_grid_shared_dg)` significant drug-gene associations across datasets, `r length(intersect(drug_grid_shared_dg, g2p_druggenes))` were G2P associations. The remaining `r length(drug_grid_shared_dg) - length(intersect(drug_grid_shared_dg, g2p_druggenes))` associations were `r drugstring2intext_converter(setdiff(drug_grid_shared_dg, g2p_druggenes)[1], "dg")` and `r drugstring2intext_converter(setdiff(drug_grid_shared_dg, g2p_druggenes)[2], "dg")`.

@@

(ref:drug-venn) **Overlap of significant drug-gene associations from CCLE, CTRP, and GDSC**. Overlap of drug-gene associations drawn from all pairwise combinations of G2P drugs and G2P genes whose drug effects significantly correlated with (**A**) mutation status, (**B**) gene expression, and (**C**) copy number across CCLE, CTRP, and GDSC.

@@

Prior to correction for an FDR of 5%, 1 DGL, `r drugstring2intext_converter(ccle_signif_g2p_lineage$Drug_Gene_Lin[1], "dgl")`, was identified in the CCLE screen when using mutation status as a correlate ($n =$ `r nrow(ccle_signif_g2p_lineage)`, `r length(unique(ccle_data_g2p_dgl$Drug_Gene_Lin)) - nrow(ccle_signif_g2p_lineage)` DGLs were not analyzed due to lack of mutation annotations in *UGT1A1* in "cancer" lineages), but CTRP ($n =$ `r nrow(ctrp_signif_g2p_lineage)`) and GDSC ($n =$ `r nrow(gdsc_signif_g2p_lineage)`) identified `r nrow(filter(ctrp_signif_g2p_lineage, !is.na(p.signif)))` and `r nrow(filter(gdsc_signif_g2p_lineage, !is.na(p.signif)))`, respectively, in skin cancer, lung cancer, and cancer lineages. After Benjamini-Hochberg FDR correction, none of the datasets yielded significant DGL interactions.

> **(???) KS test results: Need to flesh out this section**

@@

(ref:ksplot) **Comparison of drug-gene AUC vs. gene expression and vs. copy number in G2P and non-G2P cell lines**. Comparison of Spearman correlation coefficients ($r_s$) derived from copy number (log$_2$ ratio) vs. AUC z-scores and gene expression (RPKM) vs. AUC z-scores for G2P drug-gene associations. Correlation coefficients were grouped by whether or not the associated lineage for each comparison was in a level-A G2P drug-gene-lineage association or not. The distributions of these grouped coefficients were compared using Kolmogorov-Smirnov tests to identify whether correlations in G2P lineages were distinct from correlations in non-G2P lineages.

@@

```{r crispr-data, echo = FALSE, include = FALSE, warning = FALSE, message = FALSE}
crispr_data <- readRDS("./../htscreens_giant_files/crispr_data_18Q4.rds")
# Filter CRISPR for G2P genes in G2P
crispr_data_ptmuts <- filter(crispr_data, Hugo_Symbol %in% g2p_genes)
# Filter G2P CRISPR for point mutations
crispr_data_ptmuts <- filter(crispr_data_ptmuts, Reference_Allele == 1 | Reference_Allele == 0 | is.na(Reference_Allele))
crispr_data_ptmuts$Keep_PtMut_Tests <- ifelse(crispr_data_ptmuts$Reference_Allele == 1 | is.na(crispr_data_ptmuts$Reference_Allele), TRUE, ifelse(crispr_data_ptmuts$Tumor_Seq_Allele1 == 1, TRUE, FALSE))
crispr_data_ptmuts <- filter(crispr_data_ptmuts, Keep_PtMut_Tests == TRUE)
crispr_data_ptmuts <- subset(crispr_data_ptmuts, select = -c(Variant_Classification, Reference_Allele, Tumor_Seq_Allele1, Genome_Change, Keep_PtMut_Tests))
crispr_data_g2p <- unique(crispr_data_ptmuts)
crispr_data_g2p$Hugo_Symbol <- factor(crispr_data_g2p$Hugo_Symbol, levels = g2p_genes)
crispr_data_g2p$Mutation_Status_Nonsilent <- factor(crispr_data_g2p$Mutation_Status_Nonsilent, levels = c("Mutant", "Wildtype"))

@@

(ref:crispr-waterfall) Stuff stuff stuff!

```{r crispr-waterfall, echo = FALSE, message=FALSE, warning=FALSE, fig.cap='(ref:crispr-waterfall)'}
knitr::include_graphics("./plots/manuscript/crispr_spearman_corr_plot.pdf")
```

(ref:crispr-venn) **CRISPR sceen results**. Overlap of genes whose gene expression and/or copy number significantly correlated with CERES score (Spearman rank-order correlations).

@@

When we don’t account for lineage, Wilcoxon rank-sum tests (`r crispr_signif_g2p_gene_exp`, FDR < 5%) revealed that mutations in `r nrow(crispr_signif_g2p_gene_sig)` gene significantly correlated with CERES score: *`r as.character(crispr_signif_g2p_gene_sig$Hugo_Symbol)[1]`* (`r e2exp_converter(crispr_signif_g2p_gene_sig$p.adj[1], p_type = "padj")`). Gene expression (RPKM) of `r nrow(crispr_spearman_res_GE_sig)` out of `r nrow(crispr_spearman_res_GE)` (`r printnum(nrow(crispr_spearman_res_GE_sig) / nrow(crispr_spearman_res_GE) * 100)`%, Spearman rank-order correlations) G2P genes significantly correlated with CERES score, as did copy number (log2-ratio) for `r nrow(crispr_spearman_res_CN_sig)` out of `r nrow(crispr_spearman_res_CN)` (`r printnum(nrow(crispr_spearman_res_CN_sig) / nrow(crispr_spearman_res_CN) * 100)`%, Spearman rank-order correlations) G2P genes (Figure\ \@ref(fig:crispr-waterfall)). We see moderate overlap between these genes whose CERES scores significantly correlated with gene expression and with copy number (Figure\ \@ref(fig:crispr-venn)). For the purpose of comparison to the drug screen data, if we treat CRISPR knockdown like a small-molecule drug, then these results are more comprehensive than the results from CCLE, CTRP, and GDSC.

> **(???) Lineage????**

@@

In a lineage-agnostic context of the CRISPR screen, *ERBB2* gene essentiality score had significant negative correlations with *ERBB2* gene expression (`r e2exp_converter(crispr_spearman_res_GE_sig$p[which(crispr_spearman_res_GE_sig$Hugo_Symbol == "ERBB2")], "p")`, $r_s =$ `r printnum(crispr_spearman_res_GE_sig$corr[which(crispr_spearman_res_GE_sig$Hugo_Symbol == "ERBB2")])`) and *ERBB2* copy number (`r e2exp_converter(crispr_spearman_res_CN_sig$p[which(crispr_spearman_res_CN_sig$Hugo_Symbol == "ERBB2")], "p")`, $r_s =$ `r printnum(crispr_spearman_res_CN_sig$corr[which(crispr_spearman_res_CN_sig$Hugo_Symbol == "ERBB2")])`) (*Figure ##2*). Notably, gene essentiality of *MET*, which is known to have copy number alterations in some cancers [@an2014; @kim2018], did not significantly correlate with *MET* gene expression nor *MET* copy number. *BRCA1* gene essentiality also significantly correlated with both *BRCA1* gene expression (`r e2exp_converter(crispr_spearman_res_GE_sig$p[which(crispr_spearman_res_GE_sig$Hugo_Symbol == "BRCA1")], "p")`, $r_s =$ `r printnum(crispr_spearman_res_GE_sig$corr[which(crispr_spearman_res_GE_sig$Hugo_Symbol == "BRCA1")])`) and *BRCA1* copy number (`r e2exp_converter(crispr_spearman_res_CN_sig$p[which(crispr_spearman_res_CN_sig$Hugo_Symbol == "BRCA1")], "p")`, $r_s =$ `r printnum(crispr_spearman_res_CN_sig$corr[which(crispr_spearman_res_CN_sig$Hugo_Symbol == "BRCA1")])`). *BRCA2* gene essentiality significantly correlated *BRCA2* gene expression (`r e2exp_converter(crispr_spearman_res_CN_sig$p[which(crispr_spearman_res_CN_sig$Hugo_Symbol == "BRCA2")], "p")`, $r_s =$ `r printnum(crispr_spearman_res_CN_sig$corr[which(crispr_spearman_res_CN_sig$Hugo_Symbol == "BRCA2")])`) but not *BRCA2* copy number.

@@

> **(???) Should we do this?**

@@

For computing Spearman correlation coefficients, we used the `cor.test` function from the `stats` package. If there were 3 or fewer data points, then the correlation was not done and a value of `NA` was supplied.
